Reply to: 2022-05603 Intentional release of native species undermines ecological stability

### 2022-10-21 rmd/reply_r1.Rmd
Update reply_r1.Rmd

1.  **Improve the introduction and implications**: both reviewers noted an absence of context in the previous manuscript. We agree with this comment. We revised the text to contextualize our work better to address this issue. In particular, we have carefully reviewed cases to which our findings may apply and discussed the relevance of our work to those systems.
2.  **Two-species simulation.** The Editor suggested including a two-species simulation to guide the full community model. We appreciated this suggestion and were able to draw deeper insights into our theoretical analysis. This improvement helped address the concern raised by reviewer 2 (they pointed out the lack of depth in theoretical analysis). We revised the Methods and Results following this change. In particular, we removed the old sensitivity analysis as it was redundant with two-species simulations.
3.  **Ecological interpretation of our empirical results (i.e., masu salmon system).** The reviewers and editor pointed out that the interpretation of our empirical analysis is superficial -- we agree with this comment too. We have carefully enriched the text related to this issue. We have greatly increased the number of citations and discussed possible mechanisms that may produce the observed patterns more carefully. In particular, we have added a new Ricker model analysis for empirical data to estimate competition coefficients in stream fish communities. The estimated parameter indicated that competition among species certainly exists.
4.  **Life history of masu salmon.** Lastly, we enriched the text for the life history of masu salmon and how release programs for this species are operated.

We appended a point-by-point response below. We hope our manuscript is now suitable for publication in your journal.

@@

> Thank you for this constructive comment! We followed this direction, analyzing the model in a two-species community to guide a whole community analysis. As we detailed in the main text, the different combinations of $r$ and $K$ produced distinct parameter spaces where intentional release had destabilizing, neutral, and stabilizing effects. Stabilizing effects appeared when the community showed limit cycles or chaos; otherwise, the intentional release had destabilizing effects in general. We have clarified this point in the revised manuscript (see section **Theoretical prediction**).

@@

> Agreed. We have greatly improved the introduction and Discussion (paragraph L184 --, paragraph L208 --). We have clarified that our findings seem quite general for two reasons. (1) the scale of other release programs (mussels, crustaceans, birds, insects, and other fish) is generally way larger than our release program. For example, the largest release program for pink salmon in Alaska releases \> 600 million fish year^-1^ (note: our program releases only 0.24 million fish per year at the most intense site). (2) Most release programs aim to help declining populations in degraded habitats. In these habitats, the carrying capacities are likely compromised - a theoretical premise for the destabilizing effect to emerge.

>>>>

1.  **Improve the introduction and implications**: both reviewers noted an absence of context in the previous manuscript. We agree with this comment. We revised the text to contextualize our work better to address this issue. In particular, we have carefully reviewed cases to which our findings may apply and discussed the relevance of our work to those systems.
2.  **Two-species simulation.** The editor suggested including a two-species simulation to guide the full community model. We appreciated this suggestion and were able to draw deeper insights into our theoretical analysis. At the same time, this improvement helped address the concern raised by reviewer 2 (they pointed out the lack of depth in theoretical analysis). We revised the Methods and Results following this change. In particular, we removed the old sensitivity analysis as it was redundant with two-species simulations.
3.  **Ecological interpretation of our empirical results (i.e., masu salmon system).** The reviewers and editor pointed out that the interpretation of our empirical analysis is superficial; we agree with this comment. We have enriched the text pertinent to this issue. We have greatly increased the number of citations and discussed possible mechanisms (e.g., diet overlap, exploitative competition) that may produce the observed patterns more carefully. In particular, we have added a new Ricker model analysis for empirical data to estimate competition coefficients in stream fish communities. The estimated parameter indicated that competition among species certainly exists.
4.  **Life history of masu salmon.** Lastly, we enriched the text for the life history of masu salmon and how release programs for this species are operated.

In summary, the improved text/analysis provided clearer evidence that intentional release has destabilizing effects and clarified its broader relevance to other systems. We appended a point-by-point response below. We hope our manuscript is now suitable for publication in your journal.

@@

> Thank you for this constructive comment! We followed this direction, analyzing the model in a two-species community to guide a whole community analysis. As we detailed in the main text, the different combinations of $r$ and $K$ produced distinct parameter spaces where intentional release had destabilizing, neutral, and stabilizing effects. Stabilizing effects appeared when the community showed limit cycles or chaos; otherwise, the intentional release had destabilizing effects in general. We have clarified this point in the revised manuscript (see section **Theoretical prediction**; see also **Figures S1 and S2**).

@@

> Agreed. We have greatly improved the introduction and Discussion (paragraph L185 --, paragraph L209 --). We have clarified that our findings seem quite general for two reasons. (1) the scale of other release programs (mussels, crustaceans, birds, insects, and other fish) is generally way larger than our release program. For example, the largest release program for pink salmon in Alaska releases \> 600 million fish year^-1^ (note: our program releases only 0.24 million fish per year at the most intense site). (2) Most release programs aim to help declining populations in degraded habitats. In these habitats, the carrying capacities are likely compromised - a theoretical premise for the destabilizing effect to emerge.
